\boldmath Commissioning and performance in beam tests of the highly granular SiW-ECAL technological prototype for the ILC

### 2018-06-11 performance-paper-draft/draft.tex
Correction to the draft long version from Roman and myself. The intro and abstract need to be revisited

\abstract{
  High precision physics at future colliders as the International Linear Collider (ILC) require unprecedented high precision in the determination
  of the energy of final state particles.
  The needed precision will be achieved thanks to the Particle Flow algorithms (PF) which require compact, highly granular and hermetic calorimeters systems.
  The Silicon-Tungsten Electromagnetic Calorimeter (SiW-ECAL) technological prototype
  design and R\&D is tailored to the baseline design of the ECAL of the International Large Detector (ILD) for the ILC.
  In this document we present the latest news on R\&D of such prototype with fully embedded very front-end (VFE) electronics.
  Special emphasis is given to the presentation and discussion of the commissioning of the prototype
  for beam test and to the performance of the device in such beam test carried at DESY in June 2017.
}

@@

The first SiW-ECAL prototype was the so called SiW-ECAL physics prototype.
It was successfully tested at DESY, FNAL and CERN running in front of another prototype from the CALICE
collaboration, the analogue hadronic calorimeter AHCAL, delivering the proof of concept of PF calorimetry.
For the physics prototype, the VFE was placed outside the active area with no particular constraints in power consumption.
It consisted of 30 layers of Si as active material alternated with tungsten plates as absorber material.
The active layers were made of a matrix of 3x3 Si wafers. Each of these wafers was segmented in matrices of
6x6 squared pixels of 1x1 $cm^{2}$.
The prototype was divided in 3 modules of 10 layers with different W depth per layer in each of these modules
(0.4, 1.6 and 2.4 $X_{0}$) making a total of 24 $X_{0}$.
Published results proving the good performance of the technology and the PF can be found in references~\cite{Adloff:2011ha,Anduze:2008hq,Adloff:2008aa,Adloff:2010xj,CALICE:2011aa,Bilki:2014uep}.

@@

The new generation prototype is called the SiW-ECAL technological prototype and it addresses the main technological challenges: compactness,
power consumption reduction through power pulsing and VFE inside the detector close to real ILD conditions.
It will also provide data to deeply study the PF and provide input to tune the Monte Carlo programs.

@@

The sensors consist on float zone silicon wafers 320$\mu$m thick with high resistivity (larger than 5000 $\Omega\cdot$cm). The size of the wafer is $9\times9$ cm$^{2}$ featuring an array of 256 PIN diodes of $5\times5$ mm$^{2}$. A MIP traversing the PIN parallel to its normal will create $\sim$ 80 $h^{+}e^{-}$ pairs per $\mu$m which corresponds to 4.1 fC if the particle traverses the 320$\mu$m perpendicularly to its larger surface.

The original design of the silicon wafers included an edge termination made of floating guard-rings.
It was observed in beam tests \cite{Cornat:2015eoa,Cornat:2009zz} that the capacitive coupling between such floating guard-rings 
and the pixels at the edge was not negligible in tests with high energy beams (pions and electrons with energies larger than 20-40 GeV).
This coupling lead to fake events in which, at least,
the channels in the four edges of the wafer are triggered at the same time. 
This is why these events are so often called squared events. 
An R\&D program together with Hamamatsu Photonics (HPK Japan) was conducted to study the guard-rings design 
as well as the internal crosstalk. As outcome of this R\&D program, the technological prototype 
is equipped with silicon wafers with:

\begin{itemize}
\item \todo{N wafers type-X with guard-rings of XXX characteristics (continuous, segmented, segmented smaller??)}
\item \todo{N wafers type-Y without guard rings and the width of the peripheral dead areas lower than 500 $\mu$m thanks to the use of stealth dicing technique.}
\end{itemize}

For the wafers \todo{type-Y}, the amount of squared events is expected to be reduced by, at least, a \todo{factor XXX}. In both cases, 
for interaction with low energy particles as the delivered at the DESY beam test facility (see Section \ref{sec:beamtest}) the
amount of squared events is expected to be negligible, therefore we will not do any differentiation 
of our results depending on the wafer type mounted in the detector.

\todo{ Is this needed? If so, what are the right plots to show?} Before the assembly of the sensors in the detector, these are characterized one-by-one by measuring the leakage current (of the order of 1 nA per single diode) 
as a function of the bias voltage (I-V curve), the full depletion voltage (of 40 V) extracted from C-V curve and
the stability in time (leakage current versus time at a nominal bias).
%The single diodes have capacitances of 8 $pF$.

\begin{figure}[!t]
\centering
\begin{tabular}{lll}
  \includegraphics[width=1.8in]{figs/test.jpg} &
  \includegraphics[width=1.8in]{figs/test2.jpg} &
  \includegraphics[width=1.8in]{figs/test3.jpg}
\end{tabular}
\caption{I-V curve, C-V curve, leakage current vs time. \todo{Temporary picture: Triptic La Esperanza del Condenado, J. Miró}}
\label{shortslab}
\end{figure}

@@

SKIROC\cite{Callier:2011zz} (Silicon pin Kalorimeter Integrated ReadOut Chip) is the very front end ASIC (application-specific integrated circuits)
designed for the readout of the Silicon PIN didoes.
In its version SKIROC2 (and 2a) it consists of 64 channels in AMS 0.35 $\mu$m SiGe technology.
Each channel comprises a low noise charge preamplifier of variable gain followed by two lines:
a fast shaper for the trigger decision and a set of dual gain slow shaper for charge measurement. 
The gain can be controlled by modifying the feedback capacitance during the configuration of the detector.
With the slowest gain, 6pF, the ASIC will handle a linear dynamic range from 0.1 to up to 1500 MIPs 
(a slightly less than the desired final value for the ILC). 
Finally, a Wilkinson type analogue to digital converter fabricates the digitized charge deposition that can be readout. 
Once one channel is triggered, the ASIC reads out all 64 channels adding a bit of information to tag them as
triggered or not triggered and the information is stored in 15 cell deep physical switched capacitor array (SCA).
The trigger threshold can be controlled both globally and locally (channel by channel) although the
dynamical range for the channel by channel option is too small in the SKIROC version 2 due to a mistake in the development. This issue has
been solved in the SKIROC2a. The trigger holds the voltage of the two slow shapers with a tunable delay and also
holds the voltage swept with the slow clock trigger to obtain timing information.

The SKIROC ASICs can be power-pulsed by taking advantage of the ILC spill structure: 
the bias currents of the ASIC can be switched off during the idle time between bunch trains.
With this method, the ASIC is able to reduce its power consumption down to 25 $\mu$W per channel,
meeting the ILC requirements. All the results shown in this paper are obtained using the power pulsing feature.

Every ASU of the SiW-ECAL prototype is equipped with 16 SKIROCs version 2. A newer version, 2a, with some incremental improvements
and bug fixing has been produced and will be used to equip new layers currently in production.
In addition, a new generation of the ASIC, SKIROC3, is foreseen for the final detector construction.
In contrast with SKIROC2/2a, the new ASIC will be fully optimized for ILC operaton, {\it i.e.} full zero suppression, reduced power consumption etc.

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{figs/ASU_geometry1.eps}  \\
    \includegraphics[width=4in]{figs/ASU_geometry2.eps}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU. In this perspective, the Si-Sensors are in glued in the back.
    The channels are separated (in x and y) by 5.5 mm.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.
    The ASICs and channels
    are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
  }
\label{ASU}
\end{figure}

The entity
of sensors, thin PCB (printed circuit boards) and ASICs is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$ and has glued onto it 4 silicon wafers (currently with a thickness of 320 $\mu$m). The high voltage is delivered to the wafers using a HV-kapton sheet that covers the full extension of the wafers.
The ASUs are equipped
further with 16 ASIC for the read out and features 1024 square pads (64 per ASIC) of 5x5 mm.
The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU}

The readout layers of the SiW-ECAL consist of a chain of ASUs and an adapter board
to a data acquisition system (DAQ) at the beginning of the layer.
This adapter board is called SMBv4
and it also serves as a place to sit the power connectors (low voltage to power the analog and digital
parts of the ASIC and high voltage to bias the PIN diodes) and the super capacitances used for the power pulsing. 
These capacitances of 400mF with 16 m$\Omega$ of equivalent serial resistance. 
The purpose of these capacitances is to provide local storage 
of the necessary charge to avoid the transport of current pulses over long cables, 
ensuring in this way the stability of the ASICs during the acquisition.

Currently, the technological prototype layers are built with version of PCB called FEV11 with 16 SKIROC (see Section \ref{sec:DAQ}).
The FEV11 thickness is 1.6 alone and 2.7 mm including the ASICs in its current packaging: 1.1 mm thick LFBGA package.
Figure \ref{shortslab} (left) shows a picture of a full equipped short slab with FEV11 ASU.
There are ongoing R\&D activities in an alternative PCB design in which the ASICs
are directly placed on board of the PCB in dedicated cavities.
The ASICS will be in semiconductor packaging and wire bonded to the PCB. This is the so-called COB (chip-on-board) version of the ASU.
A small sample of FEV11\_COBs (same connexion pattern with the interface card than FEV11)
with a total thickness of 1.2 mm has been produced and tested in the laboratory
showing its readiness for tests with particle beams. A sample can be seen in Figure \ref{shortslab} (right).

\begin{figure}[!t]
  \centering
  \begin{tabular}{ll}
    \includegraphics[width=2.8in]{figs/test.jpg} & \includegraphics[width=2.8in]{figs/fev11_cob.png} 
  \end{tabular}
  \caption{Left \todo{Temporary picture: La Esperanza del Condenado, J. Miró}: Open single SLAB with FEV11 ASU, 16 SKIROC, interface card and DIF visibles. The silicon sensors are glued to the PCB in the other side.
  Right: two FEV11\_COB boards with 16 SKIROC2a wire bonded. The ASICs are protected with watch glasses.}
\label{ASU}
\end{figure}

@@

The design of the subsequent chain of the data acquisition (DAQ)\cite{Gastaldi:2014vaa} system is inspired by the ILC.
Current DAQ consists of three modules which are designed to be generic enough to cope with other applications.
The first module is the so called detector interface (DIF) which is placed at the beginning of each layer holding up to 15 ASUs.
All DIFs are connected by single HDMI cables to the concentrator cards: Gigabit Concentrator Card (GDCC).
The HDMI connection is used to transmit both slow control and data readout.
One GDCC controls up to 7 DIFs collecting all data from them and distributing among them the system clock and fast commands.
The most downstream module is the clock and control card (CCC).
The CCC provides a clock, control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. spill signals).

The whole system is controlled by the Calicoes and the Pyrame DAQ software version 3~\cite{Rubio-Roy:2017ere,Magniette:2018wdz}.
The Pyrame framework provides basic blocks (called modules) of control-command or data acquisition.
Calicoes is specific the implementation of these blocks for control-command and data acquisition of the SiW-ECAL prototype. 
This new version of the DAQ software implements a new mechanism to allow any module to treat and publish data in real time. 
Those data are made available to any requesting module, in particular to a new online monitor module developed
for this beam test, allowing for quick online checks of the data quality and integrity.

\subsection{Setup}
\label{sec:setup}

The prototype tested in beam is composed of 7 layers consisting in one short SLAB each.
A short SLAB is made of: one single ASU FEV11 equipped with 16 SKIROC2 and 4 wafers glued on the other side (1024 channels);
an interface card SMBv4; and a DIF. All these elements sit on a "U" shape carbon structure to protect the wafers.
The full system is then covered by two aluminum plates to provide electromagnetic shielding and mechanical stability.
The 7 layers are supported by a PVC and aluminum structure that in which can be inserted to 10 layers.
The distance between two layer slots is 15 mm and this space is used to place tungsten plates of different thicknesses.
For the beam test described in Section \ref{sec:beamtest}, the layers were placed in positions 1-6 and 10.
During the commissioning, the slabs have been tested both inside and outside of the structure.
A photograph showing the prototype can be seen in Figure \ref{proto}. 
%The mapping from this numbering scheme to real positions in the ASUs is shown in the right plot of Figure \ref{proto}.

\section{Commissioning of the detector}
\label{sec:commissioning}

This beam test was prepared by a careful and comprehensive commissioning comprising:
the debug of the
single SLABSs with special emphasis in the control of the noise and the study of the
prototype performance in cosmic rays tests.

The SKIROC2 has been already commissioned for previous experiences (see for example the 
references \cite{Amjad:2014tha,Suehara:2018mqk}). 
Therefore, all the settings used for the SKIROC2 to obtain 
the results described in this paper are motivated by these previous experiences. 
For example, the value of 1.2pF for the preamplifier. 
With this gain, the SKIROC2 ensures a linearity better than 90\% 
for 0.5-200 MIPs\footnote{\todo{Is this a fact?}}, which is enough for 
electromagnetic showers created by few GeV 
electrons or positrons. The delay of the trigger signal is the same than in 
reference \cite{Amjad:2014tha,Suehara:2018mqk}: 130 DAC for all ASICS. 
This value was chosen by selecting the delay in which an injected signal
of few MIPs of value get the maximum value of ADC. %Finally, the value of the
%compensation capacitance at the output of the preamplifier has been set to 6pF instead of 
%4pF to prevent overshots and ripples of the signal used as input to the trigger
%decision and ADC lines.

@@

The purpose of this phase was to monitor and control the
noise levels of the prototype. The commissioning includes the definition of trigger threshold values and a list of noisy channels to be masked.
Studying and control the noise levels
is crucial since noisy channels may saturate the DAQ faster than physical signals.
Two different types of noise were identified:
randomly distributed in time noisy channels 
and noise bursts affecting to all layers at the same time.

The first type of noise source (randomly distributed in time)
forced us to define a list of channels to be masked in all 
layers. Again, two different type of noisy channels have been identified. 
The first group is compose by randomly distributed in space channels.
The second group is made of
a fixed list of channels, all in the same positions for all SLABs, that were giving
underflowed ADC triggers at high rates. 
The channels associated to "underflowed-noise" events were identified by means of the online monitoring
tools. Some preliminary inspections of the FEV11 layout hint that
the cause of this "underflowed-noise" events may be in an improvable routing of the
lines from the PIN pad to the SKIROC. 
For this reason, if a given channel appeared to be associated 
to "underflowed-noise" events in one of the SLABs, 
it was masked in all of them, taking in this way the most conservative solution.
For example, all channels number 37 were masked in all ASICs.
More studies in this respect are to come.

The list of the noisy channels randomly distributed in time and space was
defined by means of dedicated data taking runs. 
These runs were characterized by: their short acquisition windows (of 2.5 ms, only 1.1 ms 
of effectively data taking time), their low
spill frequency (5 Hz) to minimize the chances of having real events due to cosmic rays 
hitting the detector during the
data taking; and the relatively high trigger threshold values between 250 and 400 DAC 
(which is equivalent to $\sim$0.5-2 MIP, see Section \ref{sec:comm_trigger} for more 
information). The full process was an iterative process starting with several repetitions
of runs at a high threshold value and then some iterations at lower threshold each up to 250 DAC. 
In every of there iterations, the channels identified as noisy were masked.
A full run involved 3-5 threshold values and at least two repetitions per value.
Again, we decided to follow a conservative approach for list the noisy channels to be masked:
if the channel was triggered at rates larger than 0.5-1\% of the total number of triggers per ASIC
it was added to the list.

In addition to the different noisy channel types described above, we also have
masked full sectors of the SLABs if an ASIC was faulty (at least 70\% of channels 
listed as noisy) or if a Si-wafer was misworking (high leakage currents). In these cases 
and in all mentioned above, the masking involved two steps: disabling of the trigger and 
the disabling of the power of the preamplifier of that channel. A summary of the results of the masking procedure is shown in Figure \ref{noisycells}.

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
  \includegraphics[width=4in]{figs/commissioning/masked_layer.eps} \\
  \includegraphics[width=4in]{figs/commissioning/masked_chip.eps}
  \end{tabular}
\caption{Ratio of channels that are marked as noisy in all slabs. 
Top: inventory of the different type of noisy channels per slab. 
Bottom: break up of the noisy channels per ASIC. We see two spikes in the number of channels masked due to routing issues in the ASICs 1 and 9. 
These two ASICs are located near the SMB where the density of lines (data and power transmission) is higher, complicating in this way the
routing. The ASICs 4-7 (wafer issue) and 10 from layer 1 and the ASIC 4 from layer 7 are not included
in the second plot since they are fully masked.}
\label{noisycells}
\end{figure}

We observed that the noise bursts happening coherently at the same time in all SLABS
was associated to cases when the electrical isolation between SLABs was broken.
This is a hint of a system effect as can be the appearence of
grounding loops or noise in the power supplies.
The issue was circumvented by improving the electrical isolation of single layers.
We have also observed that the noise bursts happened only at the end of long acquisitions therefore, in addition to the improved isolation,
we selected short enough acquisitions windows (which indeed are the most appropriate to the high rates of particles in the DESY beam).
However, more studies in the laboratory are needed in order to fully understand these issues.

@@

Once that the SKIROC main settings have been fixed and the masking of the noisy channels at a relatively high threshold value
has been done, the final step is the selection of the optimal trigger threshold values.
For that, we perform dedicated runs in which we scan a reasonable range of trigger threshold values (in DAC units)
with all channels enabled (excepted the marked as noisy). For each of the DAC points
we took data only during $\sim$ 1 minute per DAC in order to speed up the commissioning.
Plotting the total number of hits normalized to the maximum vs the DAC
for each channel gives the so-called S-curves. In the absence of external signals (cosmic rays, injected signals, etc) 
these noise S-curves show the convolution of the envelope of the 
electronic noise at the output of the fast shaper (the trigger decision line).
Indeed, the size of this enveloppe is related to the slow clock frequency
(larger periods increase the possibility to find a noise variation over the threshold).

>>>>

\abstract{
  High precision physics at future colliders as the International Linear Collider (ILC) require unprecedented high precision in the determination
  of the final state of the particles produced in the collissions
  The needed precision will be achieved thanks to the Particle Flow algorithms (PF) which require compact, highly granular and hermetic calorimeters systems.
  The Silicon-Tungsten Electromagnetic Calorimeter (SiW-ECAL) technological prototype
  design and R\&D is tailored to the baseline design of the ECAL of the International Large Detector (ILD) for the ILC.
  In this document we present and discuss the commissioning of the prototype
  and the performance of the device in a beam test carried at DESY in June 2017.
}

@@

The first SiW-ECAL prototype was the so called SiW-ECAL physics prototype.
It was successfully tested at DESY, FNAL and CERN running in front of another prototype from the CALICE
collaboration, the analogue hadronic calorimeter AHCAL, delivering the proof of concept of the technollogy
and the PF calorimetry.
For the physics prototype, the VFE was placed outside the active area with no particular constraints in power consumption.
It consisted of 30 layers of Si as active material alternated with tungsten plates as absorber material.
The active layers were made of a matrix of 3x3 Si wafers of 500 $\mu$m thickness. Each of these wafers was segmented in matrices of
6x6 squared channels of 1x1 $cm^{2}$, allowing for density of 1500 channels/dm$^{3}$
The prototype was divided in 3 modules of 10 layers with different W depth per layer in each of these modules
(0.4, 1.6 and 2.4 $X_{0}$) making a total of 24 $X_{0}$.
That very first prototype offered a signal over noise on the measured charge of 7.5 for MIP like 
particles. More results proving the good performance of the technology and the PF can be found in
references~\cite{Adloff:2011ha,Anduze:2008hq,Adloff:2008aa,Adloff:2010xj,CALICE:2011aa,Bilki:2014uep}.

@@

The new generation prototype is called the SiW-ECAL technological prototype. It addresses the main technological challenges: compactness,
power consumption reduction through power pulsing and VFE inside the detector close to real ILD conditions.
It will also provide data to deeply study the PF and provide input to tune simulation programs as for example
GEANT4\cite{Agostinelli:2002hh,Allison:2006ve,Allison:2016lfl} which is widely used
in particle physics to simulate the passage of particles through matter.

@@

The sensors consist on floating zone silicon wafers 320$\mu$m thick with high resistivity (bigger than 5000 $\Omega\cdot$cm).
The size of the wafer is $9\times9$ cm$^{2}$ and it is subdivided in an array of 256 PIN diodes of $5\times5$ mm$^{2}$.
A MIP traversing the PIN parallel to its normal will create $\sim$ 80 $h^{+}e^{-}$ pairs per $\mu$m which corresponds to 4.1 fC
for particles incident perpendicularly to its surface.

The original design of the silicon wafers included an edge termination made of floating guard-rings.
It was observed in beam tests \cite{Cornat:2015eoa,Cornat:2009zz} that the capacitive coupling between such floating guard-rings 
and the channels at the edge was not negligible in tests with high energy beams (pions and electrons with energies larger than 20-40 GeV).
This coupling lead to fake events in which, at least,
the channels in the four edges of the wafer are triggered at the same time. 
This is why these events are called squared events. 
An R\&D program together with Hamamatsu Photonics (HPK Japan) was conducted to study the guard-rings design 
as well as the internal crosstalk.
%As outcome of this R\&D program, the technological prototype 
%is equipped with silicon wafers with:
It was concluded that using wafers without guard rings and with a width of the peripheral areas lower than 
500 $\mu$m thanks to the use of stealth dicing technique, the amount of these squared events 
can be reduced to be almost negligible. This need to be confirmed in beam test.
Unfortunately, for the interaction with low energy particles as the delivered at the
DESY beam test facility (see Section \ref{sec:beamtest}) the
amount of squared events is expected to be negligible, therefore we will not discuss this issue in the following.
% do any differentiation 
%of our results depending on the wafer type mounted in the detector.

%\begin{itemize}
%\item \todo{N wafers type-X with guard-rings of XXX characteristics (continuous, segmented, segmented smaller??)}
%\item \todo{N wafers type-Y without guard rings and the width of the peripheral dead areas lower than 500 $\mu$m thanks to the use of stealth dicing technique.}
%\end{itemize}

%For the wafers \todo{type-Y}, the amount of squared events is expected to be reduced by, at least, a \todo{factor XXX}. In both cases, 
%for interaction with low energy particles as the delivered at the DESY beam test facility (see Section \ref{sec:beamtest}) the
%amount of squared events is expected to be negligible, therefore we will not do any differentiation 
%of our results depending on the wafer type mounted in the detector.

%\todo{ Is this needed? If so, what are the right plots to show?} Before the assembly of the sensors in the detector, these are characterized one-by-one by measuring the leakage current (of %the order of 1 nA per single diode) 
%as a function of the bias voltage (I-V curve), the full depletion voltage (of 40 V) extracted from C-V curve and
%the stability in time (leakage current versus time at a nominal bias).
%The single diodes have capacitances of 8 $pF$.

%\begin{figure}[!t]
%\centering
%\begin{tabular}{lll}
%  \includegraphics[width=1.8in]{figs/test.jpg} &
%  \includegraphics[width=1.8in]{figs/test2.jpg} &
%  \includegraphics[width=1.8in]{figs/test3.jpg}
%\end{tabular}
%\caption{I-V curve, C-V curve, leakage current vs time. \todo{Temporary picture: Triptic La Esperanza del Condenado, J. Miró}}
%\label{shortslab}
%\end{figure}

@@

The SKIROC\cite{Callier:2011zz} (Silicon pin Kalorimeter Integrated ReadOut Chip) is a
very front end ASIC (application-specific integrated circuits)
designed for the readout of the Silicon PIN didoes.
In its version SKIROC2 it consists of 64 channels in AMS 0.35 $\mu$m SiGe technology.
Each channel comprises a low noise charge preamplifier of variable gain followed by two branches:
a fast shaper for the trigger decision and a set of dual gain slow shaper for charge measurement. 
The gain can be controlled by modifying the feedback capacitance during the configuration of the detector.
With the slowest gain, 6pF, the ASIC will handle a linear dynamic range from 0.1 to up to 1500 MIPs 
(a slightly less than the desired final value for the ILC). 
Finally, a Wilkinson type analogue to digital converter fabricates the digitized charge deposition that can be readout. 
Once one channel is triggered, the ASIC reads out all 64 channels adding a bit of information to tag them as
triggered or not triggered and the information is stored in 15 cell deep physical switched capacitor array (SCA).

The SKIROC ASICs can be power-pulsed by taking advantage of the ILC spill structure: 
the bias currents of the ASIC can be switched off during the idle time between bunch trains.
With this method, the ASIC is able to reduce its power consumption down to 25 $\mu$W per channel,
meeting the ILC requirements. All the results shown in this paper are obtained in power pulsing mode.

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{figs/ASU_geometry1.eps}  \\
    \includegraphics[width=4in]{figs/ASU_geometry2.eps}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU. In this perspective, the Si-Sensors are in glued in the back.
    The channels are separated (in x and y) by 5.5 mm.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.
    The areas covered by the different ASICs and channels
    are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
  }
\label{ASU}
\end{figure}

The entity
of sensors, thin PCB (printed circuit boards) and ASICs is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$.
The ASUs are currently equipped
further with 16 SKIROC2 ASICs for the read out and features 1024 square pads (64 per ASIC) of 5x5 mm.
The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU}. Each ASU is equipped with 4 silicon wafers as the described in Section \ref{sec:wafers}.
The high voltage is delivered to the wafers using a HV-kapton sheet that covers the full extension of the wafers.

@@

The subsequent chain of the data acquisition (DAQ)\cite{Gastaldi:2014vaa} system is inspired by the ILC.
It consists on three modules.
The first module is the so called detector interface (DIF) which is placed at the beginning of each layer holding up to 15 ASUs;
All DIFs are connected by single HDMI cables to the concentrator cards that make
the second module: the Gigabit Concentrator Cards (GDCCs).
%The HDMI connection is used to transmit both slow control and data readout.
This cards are used to control up to 7 DIFs collecting all data from them and distributing among them the system clock and fast commands.
The last module, the most downstream, is the clock and control card (CCC) which
provides a clock, control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. signals
generated external pulse generator to simulate the ILC spill conditions).
The whole system is controlled by the Calicoes and the Pyrame DAQ software version 3~\cite{Rubio-Roy:2017ere,Magniette:2018wdz}.
%The Pyrame framework provides basic blocks (called modules) of control-command or data acquisition.
%Calicoes is specific the implementation of these blocks for control-command and data acquisition of the SiW-ECAL prototype. 
%This new version of the DAQ software implements a new mechanism to allow any module to treat and publish data in real time. 
%Those data are made available to any requesting module, in particular to a new online monitor module developed
%for this beam test, allowing for quick online checks of the data quality and integrity.

\subsection{Readout layers and SLABs}
\label{sec:setup}

\begin{figure}[!t]
  \centering
    \includegraphics[width=2.8in]{figs/test.jpg} 
  \caption{\todo{Temporary picture: La Esperanza del Condenado, J. Miró}: Open single SLAB with FEV11 ASU, 16 SKIROC 2, interface card and DIF visibles. }
\label{ASU2}
\end{figure}

The readout layers of the SiW-ECAL consist of a chain of ASUs and an adapter board
to a data acquisition system (DAQ) at the beginning of the layer.
This adapter board is called SMBv4
and it also serves as to hold other services as power connectors or the super capacitances used for the power pulsing. 
These capacitances of 400mF with 16 m$\Omega$ of equivalent serial resistance. 
The purpose of these capacitances is to provide local storage 
of the necessary charge to avoid the transport of current pulses over long cables, 
ensuring in this way the stability of the ASICs during the acquisition.

The readout layers are embedded on a "U" shape carbon structure to protect the wafers.
The full system is then covered by two aluminum plates
to provide electromagnetic shielding and mechanical stability.
This ensemble is denominated SLAB
(``short'' for 1 ASU ensembles or ``long'' for several ASUs enchained) and it can be seen in
Figure \ref{ASU2}.
With current SLABs, a potential density of
4000 channels/dm$^{3}$ is achievable. This number should be compared with
the density achieved for previous beam tests: 1500 channels/dm$^{3}$ \cite{Amjad:2014tha}.

\begin{figure}[!t]
\centering
\begin{tabular}{l}
\includegraphics[width=6.0in]{figs/assembly.png} 
\end{tabular}
\caption{Process flow for the assembly of the SiW-ECAL SLABs.}
\label{assembly}
\end{figure}

The SiW-ECAL detector designed for the ILD requires of the order
of $10^{5}$ highly integrated detection like the ones described in this text.
For the production of the small sample of SLABs studied in this document,
a scalable working procedure has been established among several groups \cite{Boudry:2318814}
profiting from the funding of projects like AIDA2020 or the HIGHTEC emblematic project
of the P2IO. A schematic view of this assembly procedure chain can be seen in
Figure \ref{assembly}. For more details we refer to Ref.\cite{Boudry:2318814}.

\section{Commissioning}
\label{sec:commissioning}

This beam test was prepared by a careful and comprehensive commissioning comprising
the debug of the
short SLABSs with special emphasis in the control of the noise and the study of the
prototype performance in cosmic rays tests.

Earlier experiences with the SKIROC2 ASIC are reported in Refs. \cite{Amjad:2014tha,Suehara:2018mqk}). 
Internal SKIROC2 parameters found in these references are adopted in the following
except if the opposite is stated.
For example, the gain value of 1.2pF for the preamplifier is used. 
With this gain, the SKIROC2 ensures a linearity better than 90\% 
for 0.5-200 MIPs, which is enough for 
electromagnetic showers created by few GeV 
electrons or positrons.
% The delay of the trigger signal is the same than in 
%reference \cite{Amjad:2014tha,Suehara:2018mqk}: 130 DAC for all ASICS. 
%This value was chosen by selecting the delay in which an injected signal
%of few MIPs of value get the maximum value of ADC. %Finally, the value of the
%compensation capacitance at the output of the preamplifier has been set to 6pF instead of 
%4pF to prevent overshots and ripples of the signal used as input to the trigger
%decision and ADC lines.

@@

%The purpose of this phase was to monitor and control the
%noise levels of the prototype.
%The commissioning includes the definition of trigger threshold values and a list of noisy channels to be masked.
Studying and control the noise levels
is crucial since noisy channels may saturate the DAQ faster than physical signals.
Two different types of noise sources were identified:
a set of noisy channels randomly distributed in time 
and noise bursts affecting to all layers at the same time.
The outcome of this commissioning is summarized in Figure \ref{noisycells}.

The first type of noise source, noise events randomly distributed in time,
forced us to define a list of channels to be masked in all 
layers. Again, two different type of noisy channels have been identified. 
The first group is compose by randomly distributed in space channels.
The second group is made of
a fixed list of channels, all in the same positions for all SLABs,
and all associated to a higher rate of events with underflowed value
of the read ADC.
%The latter set of channels were identified by means of the online monitoring
%tools.
%As channels are located in the same place for all SLABs,
%we took the conservative assumption
%that all modules would have problems in the same set of channels.
%For example, all channels number 37.
A large amount of these channels are located
in areas of the PCB where the density of lines 
(data and power transmission) is higher than in others.
This hints for an issue on the routing of the PCB.
Therefore, these channels are tagged in Figure \ref{noisycells} with the "routing issues" label
although more tests and deeper inspections of the PCB layout are needed to clarify this issue.

The list of the noisy channels randomly distributed in time and space was
defined by means of dedicated data taking runs. 
These runs were characterized by: their short acquisition windows (open to data recording for 
only 1.1 ms) at low repetition frequencies (5 Hz) to minimize the chances of having real events due to cosmic rays 
hitting the detector during the
data taking; and the relatively high trigger threshold values between 250 and 400 DAC 
(which are equivalent to $\sim$0.5-2 MIP, see Section \ref{sec:comm_trigger} for more 
information). The full process was an iterative process starting with several repetitions
of runs at a high threshold value and then some iterations at lower threshold but always 
higher than 0.5 MIP. 
In each of the iterations, the channels identified as noisy were masked.
A full run involved 3-5 threshold values and at least two repetitions per value.
Again, we decided to follow a conservative approach for list the noisy channels to be masked:
if the channel was triggered at rates larger than 0.5-1\% of the total number of triggers per ASIC
it was added to the list. These channels are labeled as "other" in the first plot in  Figure \ref{noisycells}.

In addition to the different noisy channel types described above, we also have
masked full sectors of the SLABs if an ASIC was faulty (at least 70\% of channels 
listed as noisy) or if a Si-wafer was misworking (high leakage currents). In these cases 
and in all mentioned above, the masking involved two steps: disabling of the trigger and 
the disabling of the power of the preamplifier of that channel.

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
  \includegraphics[width=4in]{figs/commissioning/masked_layer.eps} \\
  \includegraphics[width=4in]{figs/commissioning/masked_chip.eps}
  \end{tabular}
\caption{Ratio of channels that are marked as noisy in all slabs. 
Top: inventory of the different type of noisy channels per slab. 
Bottom: break down of the total number of noisy channels per ASIC. 
The ASICs 4-7 (wafer issue) and 10 from layer 1 and the ASIC 4 from layer 7 are not included
in the second plot since they are fully masked.}
\label{noisycells}
\end{figure}

The other source of noise mentioned at the beginining of this section
consist on noise bursts happening coherently at the same time in all SLABS
and it was correlated to occasions when the electrical isolation between SLABs was broken.
This is a hint of a system effect as can be the appearance of
grounding loops or disturbances in the power supplies.
The issue was circumvented by improving the electrical isolation of single layers.
We have also observed that the noise bursts happens only at the end of long acquisitions. Therefore, in addition to the improved isolation,
we selected short enough acquisitions windows (which indeed are the most appropriate to the high rates of particles in the DESY beam).
Ddedicated studies in the laboratory are currently pngoingin order to fully understand this issue.

The list of channels mentioned above is discarded from the data taking from now on.

@@

In order to select the optimal trigger threshold values for the detector operation
we perform dedicated scans of trigger threshold values
with all channels enabled (excepted the marked as noisy). 
The threshold values are given in internal DAC units which are translated to
meaningful physical quantities in Section \ref{sec:comm_trigger_sn}. 
The threshold scan curves made of the total number of hits normalized to 1 vs the threshold
for each channel are modeled by a complementary error function called ThS curves from now on:

\begin{equation}
ThS(DAC)=p_{0} \times erfc(\frac{DAC-p_{1}}{p_{2}}) = 
\newline
\frac{2p_{0}}{\sqrt(\pi)} \int_{\frac{DAC-p_{1}}{p_{2}}}^{\infty} e^{-t^{2}} dt,
\label{eq_S-curve}
\end{equation}
where $p_{0}$ is 1/2 of the normalization, $p_{1}$ is the value in which the noise levels are 
the 50\% of its maximum and $p_{2}$ give us the width of the ThS curve.

In the absence of external signals (cosmic rays, injected signals, etc) 
these noise ThS curves show the convolution of the envelope of the 
electronic noise at the output of the fast shaper (the trigger decision branch on the SKIROC).
Indeed, the size of this envelope is related to the slow clock frequency.

To reduce to the minimum the presence of cosmic rays signals, 
we perform runs with short open acquisition windows
as described in the previous section.
In Figure \ref{scurve_channels} 
two result of two threshold scans and the fit by
a ThS curves for two different channels from the second layer of the setup are shown.
